Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-06-03 ms.tex
rephrasing summary paragraph at beginning of introduction

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation and regulate the escape of gas into space. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in this process as the optical depth at line center is large. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and causing heating by collisional de-excitation. It can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy.

@@

A third class of methods are modified random walk methods \citep{1984JCoPh..54..508F,2009A&A...497..155M,Robitaille2010}. In this case a spherical volume (usually the largest within a simulation cell) is chosen when the scattering mean-free-path is small compared with the zone size.  An outgoing photon is then randomly sampled on the surface of this outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions and escape times, based on solutions to the diffusion equation withing a uniform sphere.

@@

\proposed{In this work, we present an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry and a semi-analytic solution to time-dependent \lya diffusion through the same medium. These will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.}

>>>>

The outer layers of planetary atmospheres are central to a planet's evolution, since they can shelter the lower atmosphere from high energy radiation and regulate the escape of gas into space. In atmospheres with thick layers of atomic hydrogen, stellar Lyman Alpha (Ly$\alpha$) radiation is especially important in this process as the optical depth at line center is large. Ionization and subsequent recombination can generate \lya within the planet's atmosphere, dissociating molecules in lower layers and causing heating by collisional de-excitation. It can also excite H atoms to the 2p state, creating a population of Balmer-line absorbers that can be observed via transmission spectroscopy. \todo{Need citations for these statements?}

@@

A third class of methods are modified random walk methods \citep{1984JCoPh..54..508F,2009A&A...497..155M,Robitaille2010}. In this case a spherical volume (usually the largest within a simulation cell) is chosen when the scattering mean-free-path is small compared with the zone size.  An outgoing photon is then randomly sampled on the surface of this outgoing sphere by drawing its properties from distributions in outgoing frequencies, directions and escape times, based on solutions to the diffusion equation within a uniform sphere.

@@

\proposed{The need for accurate distributions from which to draw outgoing photon properties motivates this work; namely, an analytic solution to the steady-state problem of \lya resonant scattering in spherical geometry and a semi-analytic solution to time-dependent \lya diffusion through the same medium. These will be shown to accurately represent the photon distributions in frequency and time-to-escape at the surface of a spherical volume of gas, highlighting their potential for use as a Monte Carlo accelerant in magnetohydrodynamic calculations of phenomena involving \lya resonance-line radiation.}
